---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  devtools::load_all()
)
```
# BMRSr

The goal of BMRSr is to provide wrapper functions to better make use of the Balancing Mechanism Reporting System API.

## Installation

You can install the released version of BMRSr from [GitHub](https://github.com/ARawles/BMRSr) with:

``` r
devtools::install_github("BMRSr")
```

## Example

You can use the full_request function to do a complete API request and return the results as a tibble or list

```{r full_example, eval = FALSE}
## This creates an API request to retrieve the B1720 flow on 1st Jan 2018, Settlement Period 1, and returns the results as a tibble
full_request(data_item = "B1720", api_key = "12345", settlement_date = '1 Jan 2018', period = "1", service_type = "csv")
```


You can also use the functions to create the API calls without retrieving the results

```{r build_example}
## This builds and returns the URL to retrieve the B1720 flow
build_call(data_item = "B1720", api_key = "12345", settlement_date = "1 Jan 2018", period = "1", service_type = "csv")

## This builds and returns the URL to retreive the REMIT flow
build_call(data_item = "MessageListRetrieval", api_key = "12345", publication_from = "1 Jan 2018", publication_to = "10 Jan 2018", service_type = "xml")

## This builds and returns the URL to retreive the temperature data
build_call(data_item = "TEMP", api_key = "12345", from_date = "1 Jan 2018", to_date = "10 Jan 2018", service_type = "csv")
```

